Developmental trajectories of word recognition: a priming study in bilinguals

### 2023-10-12 manuscript/manuscript.qmd
Fix data preparation and finish models

targets::tar_load(
	c(participants,
	  stimuli,
	  vocabulary,
	  attrition_trials, 
	  attrition_participants,
	  attrition_trials_vtarget,
	  attrition_participants_vtarget,
	  attrition_trials_vnone, 
	  attrition_participants_vnone,
	  attrition_trials_noeach,
	  attrition_participants_noeach,
	  bvq_data, 
	  gaze_aoi, 
	  looking_times,
	  data_time_related, 
	  data_time_cognate,
	  data_time_related_vtarget,
	  data_time_cognate_vtarget,
	  data_time_related_vnone,
	  data_time_cognate_vnone,
	  data_time_related_noeach, 
	  data_time_cognate_noeach))

targets::tar_load(
	c(model_fits_related,
	  model_fits_cognate,
	  model_fits_related_vtarget,
	  model_fits_cognate_vtarget,
	  model_fits_related_vnone,
	  model_fits_cognate_vnone,
	  model_fits_related_noeach,
	  model_fits_cognate_noeach,
	  model_loo_related,
	  model_loo_cognate,
	  model_loo_related_vtarget,
	  model_loo_cognate_vtarget,
	  model_loo_related_vnone,
	  model_loo_cognate_vnone,
	  model_loo_related_noeach,
	  model_loo_cognate_noeach
	))

@@

>>>>

targets::tar_load(
	c(participants,
	  stimuli,
	  vocabulary,
	  attrition_trials, 
	  attrition_participants,
	  attrition_trials_vtarget,
	  attrition_participants_vtarget,
	  attrition_trials_vnone, 
	  attrition_participants_vnone,
	  attrition_trials_noeach,
	  attrition_participants_noeach,
	  attrition_trials_vnoeach,
	  attrition_participants_vnoeach,
	  bvq_data, 
	  gaze_aoi, 
	  looking_times,
	  data_time_related, 
	  data_time_cognate,
	  data_time_related_vtarget,
	  data_time_cognate_vtarget,
	  data_time_related_vnone,
	  data_time_cognate_vnone,
	  data_time_related_noeach, 
	  data_time_cognate_noeach,
	  data_time_related_vnoeach,
	  data_time_cognate_vnoeach))

targets::tar_load(
	c(model_fits_related,
	  model_fits_cognate,
	  model_fits_related_vtarget,
	  model_fits_cognate_vtarget,
	  model_fits_related_vnone,
	  model_fits_cognate_vnone,
	  model_fits_related_noeach,
	  model_fits_cognate_noeach,
	  model_fits_related_vnoeach,
	  model_fits_cognate_vnoeach,
	  model_loo_related,
	  model_loo_cognate,
	  model_loo_related_vtarget,
	  model_loo_cognate_vtarget,
	  model_loo_related_vnone,
	  model_loo_cognate_vnone,
	  model_loo_related_noeach,
	  model_loo_cognate_noeach,
	  model_loo_related_vnoeach,
	  model_loo_cognate_vnoeach
	))

@@

## Analysis 5

Participants need to know the prime or target words, and do not need to look to both pictures.

```{r dataset-vnoeach}
n_trials <- nrow(attrition_trials_vnoeach)
n_trials_valid <- inner_join(attrition_participants_vnoeach,
							 attrition_trials_vnoeach) |> 
	filter(is_valid_participant) |> 
	pull(is_valid_trial) |> 
	sum()
n_participants_valid <- sum(attrition_participants_vnoeach$is_valid_participant)
n_exc_prime <- sum(!attrition_trials_vnoeach$is_valid_gaze_prime)
n_exc_test <- sum(!attrition_trials_vnoeach$is_valid_gaze_test)
n_exc_test_each <- sum(!attrition_trials_vnoeach$is_valid_gaze_test_each)
n_exc_vocab <- sum(!attrition_trials_vnoeach$is_valid_vocab)
n_exc_cognate <- sum(!attrition_participants_vnoeach$is_valid_cognate)
n_exc_noncognate <- sum(!attrition_participants_vnoeach$is_valid_noncognate)
n_exc_unrelated <- sum(!attrition_participants_vnoeach$is_valid_unrelated)

n_longitudinal <- attrition_participants_vnoeach |>
	filter(is_valid_participant) |>
	count(id, name = "times") |> 
	count(times)
```

```{r tbl-attrition-trials-vnoeach}
#| label: tbl-attrition-trials-vnoeach
#| tbl-cap: "Trial attrition rate by condition for included participants. Additional excluded trials are indicated between parentheses."
attrition_trials_vnoeach |> 
	filter(id %in% attrition_participants_vnoeach$id[attrition_participants_vnoeach$is_valid_participant]) |> 
	left_join(select(participants, filename, id, age_group),
			  by = join_by(filename, id, age_group)) |> 
	summarise(n_valid = sum(is_valid_trial),
			  n_total = n(),
			  .by = c(id, age_group, trial_type)) |> 
	summarise(across(n_valid, lst(sum, mean, sd),
					 .names = "{.fn}"),
			  n_total = sum(n_total),
			  .by = c(age_group, trial_type)) |> 
	mutate(n_excluded = n_total-sum) |> 
	select(-c(n_total)) |> 
	pivot_wider(names_from = trial_type,
				values_from = c(sum:sd, n_excluded),
				names_repair = janitor::make_clean_names) |> 
	rename_with(\(x) gsub("non_cognate", 
						  "noncognate",
						  x)) |> 
	arrange(age_group) |> 
	relocate(age_group,
			 matches("_cognate"),
			 matches("noncognate")) |> 
	gt(rowname_col = "age_group") |> 
	grand_summary_rows(columns = matches("mean_"),
					   fns = lst(Mean ~ mean(.)),
					   fmt = ~fmt_number(.)) |>
	grand_summary_rows(columns = matches("sum_"),
					   fns = lst(Sum ~ sum(.)),
					   fmt = ~fmt_integer(.)) |>
	cols_merge(c(sum_cognate, n_excluded_cognate), 
			   pattern = "{1} ({2})") |> 
	cols_merge(c(sum_noncognate, n_excluded_noncognate), 
			   pattern = "{1} ({2})") |> 
	cols_merge(c(sum_unrelated, n_excluded_unrelated),
			   pattern = "{1} ({2})") |> 
	cols_merge_uncert(mean_cognate, sd_cognate) |> 
	cols_merge_uncert(mean_noncognate, sd_noncognate) |> 
	cols_merge_uncert(mean_unrelated, sd_unrelated) |> 
	tab_spanner("Cognate trials", ends_with("_cognate")) |>
	tab_spanner("Non-cognate trials", ends_with("noncognate")) |> 
	tab_spanner("Unrelated trials", ends_with("unrelated")) |> 
	tab_spanner("Related trials", matches("cognate")) |>
	fmt_number(matches("mean|sd")) |> 
	fmt_integer(matches("sum"), sep_mark = ",") |> 
	cols_label(sum_cognate = "N",
			   sum_noncognate = "N",
			   sum_unrelated = "N",
			   mean_cognate = "Mean",
			   mean_noncognate = "Mean",
			   mean_unrelated = "Mean")
```

A model including the *Relatedness* $\times$ *Group* interaction showed the best of-of-sample predictive performance, although the model including only *Relatedness* performed equivalently ($\text{ELPD}_{\mathcal{M_0}} - \text{ELPD}_{\mathcal{M_1}}$ = `r round(model_loo_related_vnoeach[2, 1], 3)`, *SE* = `r round(model_loo_related_vnoeach[2, 2], 3)`). Both models showed substantially better predictive performance than the model including only *Group* ($\text{ELPD}_{\mathcal{M_0}} - \text{ELPD}_{\mathcal{M_2}}$ = `r round(model_loo_related_vnoeach[3, 1], 3)`, *SE* = `r round(model_loo_related[3, 2], 3)`). This indicates that including the *Relatedness* predictor improved the predictive performance of the model significantly, that including its interaction with *Group* slightly increased the performance of the model, and that the main effect of *Group* by itself barely changed the predictive performance of the model.

```{r fig-related-vnoeach}
#| label: fig-related-vnoeach
#| fig-height: 6
#| fig-width: 8
#| fig-cap: "Marginal posterior predictions of the GAMMs. (A) Mean posterior probability of target looking across the time course of the trial. Black lines and intervals indicate the psoterior mean and 95% credible intervals. Points indicate the mean probability of target looking across participants. (B) Difference in posterior probability of target looking between *related* and *unrelated* trials. The yellow rectangle indicates, in both A and B, the range of time points in which the 95% credible interval of the differences excluded zero."
epreds <- expand_grid(condition = levels(data_time_related_vnoeach$condition),
					  timebin = seq(0, 17, length.out = 100),
					  age = mean(data_time_related_vnoeach$age),
					  lp = levels(data_time_related_vnoeach$lp),
					  .nsamples = 1) |>
	add_epred_draws(model_fits_related_vnoeach[[4]],
					ndraws = NULL,
					re_formula = NA, 
					value = ".value") |> 
	mutate(lp = factor(lp, levels = c("Monolingual", "Bilingual")))

data_time_related_vnoeach |> 
	summarise(.prop = mean(.prop),
			  .by = c(id, timebin, lp, condition, age)) |> 
	ggplot(aes(timebin, .prop, 
			   colour = condition,
			   fill = condition,
			   shape = condition,
			   linetype = condition)) +
	facet_wrap(~lp) +
	# geom_rect(data = diff_rect,
	# 		  aes(xmin = xmin,
	# 		  	xmax = xmax,
	# 		  	ymin = -Inf,
	# 		  	ymax = Inf),
	# 		  colour = NA,
	# 		  fill = "orange",
	# 		  alpha = 1/2,
	# 		  inherit.aes = FALSE) +
	# geom_line(data = epreds,
	# 		  aes(y = .epred,
# 		  	group = interaction(condition, .draw)),
# 		  linetype = "solid",
# 		  alpha = 0.1,
# 		  linewidth = 3/4) +
stat_summary(data = epreds,
			 aes(y = .value),
			 fun.data = \(x) mean_qi(x, .width = 0.95),
			 geom = "ribbon",
			 alpha = 0.5,
			 linewidth = 0) +
	stat_summary(data = epreds,
				 aes(y = .value,
				 	linetype = condition),
				 fun = "mean",
				 geom = "line",
				 colour = "black",
				 linewidth = 3/4) +
	geom_hline(yintercept = 1/2, 
			   linewidth = 1/2,
			   colour = "black",
			   linetype = "dotted") +
	stat_summary(fun = mean,
				 geom = "point",
				 colour = "black",
				 size = 2.5,
				 stroke = 3/4) +
	labs(x = "Time (ms)",
		 y = "P(Target looking)",
		 colour = "Condition",
		 fill = "Condition",
		 linetype = "Condition",
		 shape = "Condition") +
	theme(legend.title = element_blank(),
		  axis.title.x = element_blank()) +

```{r fig-cognate-vnoeach}
#| label: fig-cognate-vnoeach
#| fig-height: 6
#| fig-width: 8
#| fig-cap: "Marginal posterior predictions of the GAMMs. (A) Mean posterior probability of target looking across the time course of the trial. Black lines and intervals indicate the psoterior mean and 95% credible intervals. Points indicate the mean probability of target looking across participants. (B) Difference in posterior probability of target looking between *cognate* and *non-cognate* trials. The yellow rectangle indicates, in both A and B, the range of time points in which the 95% credible interval of the differences excluded zero."
epreds <- expand_grid(condition = levels(data_time_cognate_vnoeach$condition),
					  timebin = seq(0, 17, length.out = 100),
					  age = mean(data_time_cognate_vnoeach$age),
					  lp = levels(data_time_cognate_vnoeach$lp),
					  .nsamples = 1) |>
	add_epred_draws(model_fits_cognate_vnoeach[[4]],
					ndraws = NULL,
					re_formula = NA) |> 
	mutate(lp = factor(lp, levels = c("Monolingual", "Bilingual")))

data_time_cognate_vnoeach |> 
	summarise(.prop = mean(.prop),
			  .by = c(id, timebin, lp, condition, age)) |> 
	ggplot(aes(timebin, .prop, 
			   colour = condition,
			   fill = condition,
			   shape = condition)) +
	facet_wrap(~lp) +
	# geom_rect(data = diff_rect,
	# 		  aes(xmin = xmin,
	# 		  	xmax = xmax,
	# 		  	ymin = -1.5,
	# 		  	ymax = 1.5),
	# 		  colour = NA,
	# 		  fill = "orange",
	# 		  alpha = 1/2,
	# 		  inherit.aes = FALSE) +
	# geom_line(data = epreds,
	# 		  aes(y = .epred,
# 		  	group = interaction(condition, .draw)),
# 		  linetype = "solid",
# 		  alpha = 0.1,
# 		  linewidth = 3/4) +
stat_summary(data = epreds,
			 aes(y = .epred),
			 fun.data = \(x) mean_qi(x, .width = 0.95),
			 geom = "ribbon",
			 alpha = 0.5,
			 linewidth = 0) +
	stat_summary(data = epreds,
				 aes(y = .epred,
				 	linetype = condition),
				 fun = "mean",
				 geom = "line",
				 colour = "black",
				 linewidth = 3/4) +
	geom_hline(yintercept = 0.5, 
			   linewidth = 1/2,
			   colour = "black",
			   linetype = "dotted") +
	stat_summary(fun = mean,
				 geom = "point",
				 colour = "black",
				 size = 2.5,
				 stroke = 3/4) +
	labs(x = "Time (ms)",
		 y = "P(Target looking)",
		 colour = "Prime type",
		 fill = "Prime type",
		 linetype = "Prime type",
		 shape = "Prime type") +
	theme(legend.title = element_blank(),
		  axis.title.x = element_blank()) +
